\documentclass{article}

\newcommand{\doctitle}{Assignment 4} % change document title
\usepackage{ellis_header}

\begin{document}

\section{Chapter 2.1}
\setcounter{problem}{3}
\begin{problem} % Problem 4
Consider the possible functions $f: [7] \rightarrow [9]$
\begin{enumerate}[(a)]
    \item How many have $f(3) = 8$? 
    \item How many have $f(3) \neq 8$?
    \item How many have $f(1) \neq 5$ and are one-to-one?
    \item How many have $f(i)$ even, for all $i$?
    \item How many have $\text{rng}(f) = \{5, 6 \}$?
    \item How many in which $f^{-1}$ is not a function?
\end{enumerate}
\end{problem}

\textbf{Answer:}
\begin{enumerate}[(a)]
    \item $f(3)$ has 1 choice. This leaves 6 other values that can go into 9 other bins. I have nine bins, that I can place six values in, with repitition allowed. This is $9^6$.
    \item $f(3)$ has 8 choices, while the six other values have 9 choices. This is $8 \cdot 9^6$.
    \item $f(1)$ has 8 choices, $f(2)$ has eight choices (since one was removed by $f(1)$.) Then $f(3)$ has seven choices, $f(4)$ has six choices... This is $8 \cdot (8)_6 = \dfrac{8 \cdot 8!}{(8-6)!}$.
    \item Instead of 9 choices, we now have 4 choices. So 7 values have 4 choices. This is $4^7$.
    \item Now we have only two choices, $\{5, 6\}$ for 7 values. This is $2^7$. But, we must guarentee at least one values in $\{5\}$ and one value in $\{6\}$. So subtract out when all the values map to 5 or all the values map to 6, which is two cases. Final answer is $2^7 - 2$.
    \item For $f^{-1}$ to be a function, the original function $f$ must be one-to-one. We are looking for where the original function is not one-to-one. Base options is 9 choices for 7 values. So $9^7$. But, we need to subtract out the one-to-one option, which is $(9)_7$. The final answer is $9^7 - (9)_7$.
\end{enumerate}

\setcounter{problem}{6}
\begin{problem} % Problem 7
Find the number of onto functions $[k] \rightarrow [4]$.
\end{problem}

\textbf{Answer:} I am assuming that $k \ge 4$. The base case is $4^k$ options for distribution. Then need to subtract out the following:
\begin{enumerate}[(a)]
    \item Missing 3 of the elements in $[4]$. Answer is $4^1 = 4$.
    \item Missing 2 of the elements in $[4]$. You have $\binom{4}{2} = 6$ elements to miss. You then have $2^k$ ways to distribute the values to the remaining elements. Each distribution has 2 ways that all the values end up on one elements instead of both, so remove those. The answer is $6(2^k - 2)$.
    \item Missing 1 of the elements in $[4]$. You have $\binom{4}{3}= 4$ different elements to miss. Once you miss, you can select $3^k$ places to go. But you need to guarentee that you don't miss more than 1 element. There are 3 ways to miss 2 elements with $2^k$ places to go. There are 3 ways to miss 3 elements with $1^k$ places to go. The answer is $4(3^k - 3 (2^k - 2) - 3 \cdot 1^k)$
\end{enumerate}

Final answer is $4^k - (4 + 6(2^k -2) + 4(3^k -(3 + 3( 2^k - 2))))$

\setcounter{problem}{10}
\begin{problem} % Problem 11
Give a combinatorial proof. For $n \ge 1$ and $k \ge 1$, then $2^{kn} > \max \{n^k, k^n\}$. Hint: Compare relations to functions.
\end{problem}

\textbf{Answer:} Before diving into the details with the general case, I'm going to do a numerical example to get us warmed up. Let us define two sets.
\begin{equation*}
    K = [4] = \{1, 2, 3, 4\} \quad N = [3] = \{1, 2, 3\}
\end{equation*}

All the possible relations from $K$ to $N$ are contained in the Cartesian product of $K \times N$.
\begin{equation}
    K \times N = \{\{1,1\}, \{1, 2\}, \{1, 3\} \dots \{4, 1\}, \{4, 2\}, \{4, 3\}\}
\end{equation}

Every possible relation on $K \rightarrow N$ is a subset of $K \times N$. For instance, for the relation $>$ would have $\{\{2, 1\}, \{3, 1\}, \{3, 2\}, \{4, 1\}, \{4, 2\}, \{4, 3\}\}$. So all the possible relations is $\mathscr{P}(K \times N)$, the power set of the cartesian product of $K \times N$. So, the number of possible relations is $2^{|K \times N|}$. If $k = |K|$ and $n = |N|$ then it is $2^{kn}$.

\medskip
Now, we focus our attention to functions instead of relations. The basic defintion of a function is that every input can only have one output. Thinking about this as distributions, the first number in $K$ has 3 potential choices. The second number in $K$ also has 3 choices, and so one. As a result, when looking at functions, there are $n^k = 3^4 = 81$ possible functions from $K \rightarrow N$. Likewise, if we go the opposite way, from $N \rightarrow K$, there are $k^n = 4^3 = 64$ possible functions.

\medskip
What this is saying, is the number of possible relations between two sets will always be bigger than the possible functions from one set to another and vice a versa.

\begin{proof}
Suppose that $k, n \ge 1$ and $K,N$ are sets of the first postive integers $k,n$ denoted as $K = [k]$ and $N = [n]$. Then the set of all possible relations between $K$ and $N$ is the powerset of $K \times N$, denoted as $\mathscr{P}(K \times N)$. The cardinality of the powerset is $|\mathscr{P}(K \times N)| = 2^{kn}$, or the number of possible relations.

\medskip
For a relation to be considered a function, every value in the domain must map to only one value in the codomain. The function $f: K \rightarrow N$ has $n^k$ possible variations. The function variations is a subset of the relation variations, written $f \subseteq \mathscr{P}(K \times N)$. As such, $n^k \le 2^{nk}$. This inequality can be written $n^k < 2^{nk}$ by recognizing the empty set is inside the powerset but will never be considered to meet the definition of a function.

\medskip
WLOG: Repeat the functional arguement for $f: N \rightarrow K$ with $k^n$ possible variations. This proves that $2^{kn} > \max \{n^k, k^n\}$.
\end{proof}

\section{Chapter 2.2}
\setcounter{problem}{3}
\begin{problem} % Problem 4
Only include a, b, d
\end{problem}

\begin{problem} % Problem 5
Rawr
\end{problem}

\setcounter{problem}{6}
\begin{problem} % Problem 7
Only include a, b, c
\end{problem}

\section{Chapter 2.3}
\setcounter{problem}{1}
\begin{problem} % Problem 2
Rawr
\end{problem}


\setcounter{problem}{6}
\begin{problem} % Problem 7
Rawr
\end{problem}


\end{document}